---
title: "R package development"
author: "Jae Yeon Kim"
institute: "UC Berkeley"
date: "`r Sys.Date()`"
---

# Objectives

[Shiny](https://shiny.rstudio.com/) is a "framework for creating web applications using R code". You can create a dashboard or an interactive map without knowing anything about HTML, CSS, or JavaScript. Developing a shiny app helps people with little technical expertise to learn from your data intuitively and interactively.

```{=html}

<iframe width="560" height="315" src="https://www.youtube.com/embed/Wy3TY0gOmJw" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

<p>Shiny in production: Principles, practices, and tools - Joe Cheng</p>

```

```{=html}

<iframe src="https://vac-lshtm.shinyapps.io/ncov_tracker/?_ga=2.240702211.1091983227.1603295793-100003412.1602392815" style="border: 1px solid #AAA; width: 800px; height: 700px"></iframe>

<p>COVID-19 tracker by Edward Parker</p>

```

# Workflow 

The workflow follows what Hadley Wickham recommended in his book on mastering shiny. 

1. Install libraries 

```r
install.packages("shiny")
```

2. Create app directory and file 

Add an `app.R` file.

The key objective here is defining your UI (how the app looks; front-end = INPUT) (defined in object `ui`) and server (how the app works; back-end = OUTPUT) (defined in object `server`).

If you're creating a complex app, you can achieve the same goal with two files: `ui.R` and `server.R`.

### app.r 

- Front-end

```{r eval=FALSE}

# Load packages 
# Do not use install.packages(), pacman::p_load(), or library() if you intend to deploy the app using shinyapps.io 

require("wordcloud2")
require("shiny")
require("shinydashboard")
require("colourpicker")

# Load data 

df <- read.csv(url("https://github.com/jaeyk/covid19antiasian/raw/master/processed_data/hash_counts.csv"))[,-1]

# Defines the user interface; how the app looks

ui <- fluidPage(
  
    # Application title 
    titlePanel("Word Cloud on the Hashtags of the Tweets related to COVID-19 & Asian|Chinese|Wuhan"),
  
    h4(tags$a(href = "https://jaeyk.github.io/", "Developer: Jae Yeon Kim")),
            
    sidebarLayout(
      
      # Sidebar with sliders 
      sidebarPanel(
        sliderInput("size", 
                    "Font size:",
                    min = 1, max = 10,
                    value = 2)
      ),
    
    mainPanel(
          
          wordcloud2Output("cloud"),
        
        )
    
    )
)
  
```

- Back-end

```{r eval=FALSE}
server <- function(input, output, session) {
  
  output$cloud <- renderWordcloud2({ 
    
    wordcloud2(df, 
               size = input$size, 
               color = "random-dark") 
    
    })

  }
```

- Build a shiny app 

```{r eval=FALSE}
shinyApp(ui = ui, server = server)
```

# Deployment 

- Deploy to [the shinyapps.io cloud](https://www.shinyapps.io/?_ga=2.5503866.871102833.1602978469-100003412.1602392815) 

```{r eval = FALSE}

# Install packages 
install.packages("rsconnect")
library(rsconnect)

# Setup 
rsconnect::setAccountInfo(name = "<Account name>", 
                          token = "<Token>",
                          secret = "<Secret>")

rsconnect::deployApp(appNames = "<App name>")
```

# References 

[Mastering Shiny](https://mastering-shiny.org/) by Hadley Wickham. For newbies. 

[Shiny Documents](https://bookdown.org/yihui/rmarkdown/shiny-documents.html) by Yihui Xie

[Engineering Production-Grade Shiny Apps](https://engineering-shiny.org/) by Colin Fay, Sébastien Rochette, Vincent Guyader, Cervan Girard.

[Building Shiny Apps](https://stat545.com/shiny-tutorial.html) by Dean Attali.

## Other useful data products 

- Automating data reports using rmarkdown (called [parameterized reports](https://rmarkdown.rstudio.com/developer_parameterized_reports.html%23parameter_types%2F))
- Automating R presentation using [slidify](http://slidify.org/index.html)
- Creating interactive web apps using [leaflet](https://rstudio.github.io/leaflet/) 
