---
title: "Flow Control"
subtitle: ""
author: "Jae Yeon Kim"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

> Anything that can be automated should be automated. Do as little as possible by hand. Do as much as possible with functions. - Hadley Wickham 

# Big picture

This week's lecture and section helps you to step up your R skills with functional programming. The `purrr` package provides easy-to-use tools to automate repeated things in your entire R workflow (e.g., wrangling, modeling, and visualization). The end result is cleaner, faster, more readable and extendable code.

![](https://encrypted-tbn0.gstatic.com/images?q=tbn:ANd9GcSmywiiOutD0NPieYCKxaD2wN9Fbt2I3iS87A&usqp=CAU)

## Learning objectives

0.  How to use control flow in R using `if_`, `for loop`, and `apply`
1.  How to use `map()` to automate workflow in a cleaner, faster, and more extendable way
2.  How to use `map2()` and `pmap()` to avoid writing nested loops 
3.  How to use `map()` and `glue()` to automate creating multiple plots
4.  How to use `reduce()` to automate joining multiple dataframes
5.  How to use `slowly()` and `future_` to make automation process either slower or faster
6.  How to use `safely()` and `possibly()` to make error handling easier

# Setup

```{r}
# Install packages
if (!require("pacman")) {
  install.packages("pacman")
}

pacman::p_load(
  tidyverse, # tidyverse pkgs including purrr
  bench, # performance test 
  tictoc, # performance test
  broom, # tidy modeling
  glue, # paste string and objects
  furrr, # parallel processing
  rvest, # web scraping
  devtools, # dev tools 
  usethis, # workflow     
  roxygen2, # documentation 
  testthat, # testing 
  patchwork, # arranging ggplots
  purrr)
```

## Flow control

-   Control structures = putting logic in code to control flow (e.g., `if`, `else`, `for`, `while`, `repeat`, `break`, `next`)

-   Almost all the conditional operators used in Python also work in R. T
he basic loop set up is also very similar, with some small syntax adjustments.

-   `if()` is a function whose arguments must be specified inside parentheses.

```{r eval = FALSE}
# This is pseudo code so it should not run 
if (my_kid_behavior == cry){
  
  print("I will sing Baby Shark.")
  
}
```

-   `else`, however, is a reserved operator that takes no arguments. Note that there is no `elif` option --- one simply writes `else if()`.

```{r eval = FALSE}

# Two conditions 

if (my_kid_behavior == cry){
  
  print("I will sing Baby Shark.")

} else {print("I will take a nap.")}

# Three conditions 

if (my_kid_behavior == cry){
  
  print("I will sing Baby Shark.")

}

else if (my_kid_behavior == smile){
  
  print("I will also smile.")
  
} else {print("I will take a nap.")}

```

-   Whereas operations to be executed after conditional evaluations in Python come after a `:`, R operations must only be enclosed in curly brackets: `{}`. Furthermore, there is no requirement for indentation.

### if (one condition)

```{r}
x <- 5

if (x < 0) { # Condition 
  print("x is negative") # Do something 
} 
```

```{r}
x <- -5

if (x < 0) {
  print("x is negative")
}
```


### if + else (two conditions)

```{r}
x <- 5

if (x < 0) {
  print("x is negative")
} else{
  print("x is positive")
}
```

### if + else if + else (three conditions)

```{r}
x <- 0

if (x < 0) { # Condition 
  print("x is negative") # Do something 

  } else if (x == 0) { 
  
    print("x is zero") # Do something else 

    } else {print("x is positive") # Do something else 

      }
```

-   In general, it's not a good idea to write nested code (lots of `else_if()` or `ifelse()`). It's difficult to read, debug, modulize, and extend.
-   Instead, write functions and, if necessary, use `if()` only. We'll come back to this later.

### Functions

While functions are defined in Python using the `def` reserved operator, R sees functions as just another type of named object. Thus, they require explicit assignment to an object. This is done using the function `function()`, which creates a function taking the arguments specified in parentheses.

function = input + computation (begin -\> end) + output

```{r}
simple.function <- function(x){ # x = input
  print(x + 1) # computation as well as output 
}
```


```{r}
simple.function(x = 2)

less.simple.function <- function(x, y){
  print(x - y + 1)
}

less.simple.function(x = 2, y = 10)
```

With respect to returning function output, most of the same rules apply as with Python. Be sure to remember that `return()` will only process a single object, so multiple items must usually be returned as a list. Note that your ordering of the functions matters, too.

```{r}
dumbfun <- function(x){
  return(x)
  print("This will never print :(")
}

dumbfun(x = "something")

dumbfun <- function(x){
  print("Why did I print?")
  return(x)
}

dumbfun(x = "something")

dumbfun <- function(x,y){
  thing1 <- x
  thing2 <- y
  return(list(thing1, thing2))
}

dumbfun(x = "some text", y = "some data")
dumbfun(x = c(5,10,15), y = "some data")
```

R functions also allow you to set default argument values:

```{r}
less.simple.function <- function(x, y = 0){
  print(x - y + 1)
}

less.simple.function(x = 2)

less.simple.function(x = 2, y = 10)
```

With respect to specifying arguments, one can either use argument **position** specifications (i.e., the order) or argument **name** specifications. The latter is strongly preferred, as it is very easy to accidentally specify incorrect argument values.

```{r}
send <- function(message, recipient, cc = NULL, bcc = NULL){
  print(paste(message, recipient, sep = ", "))
  print(paste("CC:", cc, sep = " "))
  print(paste("BCC:", bcc, sep = " "))
}

send(message = "Hello", recipient = "World", cc = "Sun", bcc = "Jane")

send("Hello", "World", "Sun", "Jane")

send("Hello", "Sun", "Jane", "World")

send(message = "Hello", cc = "Sun", bcc = c("Jane", "Rochelle"), recipient = "World")
```

Also, note that functions don't have what CS people called side-effects. Functions only define local variables = They don't change objects stored in the global environment. (Consider the difference between `<-` and `=` for assignments.) That's why you can use functions for reusable tasks since it does not interrupt other important things in your system.

See [the following example](https://darrenjw.wordpress.com/2011/11/23/lexical-scope-and-function-closures-in-r/) from Wilkinson.

```{r}
a = 1 
b = 2

f <- function(x)
{
  a*x + b
}

f(2)

g <- function(x)
{
  a = 2
  b = 1
  f(x)
}

g(2) # a equals still 1 
```

**Additional tips**

-   Nonstandard evaluation

Nonstandard evaluation is an advanced subject. If you feel overwhelmed, you are more than welcome to skip this. But if you are serious in R programming, this is something you want to check out. For deeper understanding of this issue, I recommend reading [Ren Kun's very informative blog post](https://renkun.me/2014/12/03/tips-on-non-standard-evaluation-in-r/) carefully.

This part draws on one of the [the dplyr package articles](<https://dplyr.tidyverse.org/articles/programming.html>.

In tidyverse, calling a variable with or without quotation mark (string or not) does make little difference because tidyeval is a type of non-standard evaluation. This flexibility runs into the following problem when it comes to programming.

```{r eval = FALSE}

# Using `mpg` instead of `mtcars$mpg` is called data masking.

mtcars %>% select(mpg)

mtcars %>% select("mpg")

```

Data and env-variables

```{r}
# df = environment variable 
df <- data.frame(
  x = c(1:5),
  y = c(6:10)
  )

# x, y = data variables 
df$x
df$y
```

-   Problem

```{r}
x <- NULL 

var_summary <- function(env_var, data_var){
 
   env_var %>%
    summarise(mean = mean(data_var))

}
```

You may expect that the output is mean = 2.5 ... but

It's because the mean() function doesn't take `df$x` for data_var but `x`. You need to link x with environment variable

```{r}

var_summary(df, x)

```

This is how you can fix this.

```{r}

# Solution
vs_fix <- function(env_var, data_var){
 
   env_var %>%
    summarise(mean = mean({{data_var}}))

}

# You can also do this. 
vs_fix_enhanced <- function(env_var, data_var){
 
   env_var %>%
    summarise("mean_{{data_var}}" := mean({{data_var}})) # If you use the glue package, this syntax is very intuitive.

}

vs_fix_enhanced(df, x)

```

If you have a character vector input ...

```{r}
mtcars_count <- mtcars %>%
  names() %>%
  purrr::map(~count(mtcars, .data[[.x]])) # We're going to learn about map in the rest of this session.

mtcars_count[[1]]
```

### for loop

![Concept map for a for loop. Source: <https://teachtogether.tech/en/index.html#s:memory-concept-maps>](https://teachtogether.tech/en/figures/for-loop.svg)

Loops in R also work basically the same way as in Python, with just a few adjustments. First, recall that index positions in R start at 1. Second, `while()` and `for()` are functions rather than reserved operators, meaning they must take arguments in parentheses. Third, just like `else`, the `in` operator *is* reserved and takes no arguments in parentheses. Fourth, the conditional execution must appear between curly brackets. Finally, indentation is meaningless, but each new operation must appear on a new line.

-   `while()`: when we have no idea how many times loop needs to be executed.
-   `for()`: when we know how many times loop needs to be executed. This is likely to be the loop you are going to use most frequently.

```{r}
fruits <- c("apples", "oranges", "pears", "bananas")

# a while loop
i <- 1

while (i <= length(fruits)) {
  print(fruits[i])
  i <- i + 1
}

# a for loop
for (i in 1:length(fruits)) {
  print(fruits[i])
}
```

### apply family

While and for loops in R can be very slow. For this reason, R has a number of built-in iteration methods to speed up execution times. In many cases, packages will have "behind-the-scenes" ways to avoid for loops, but what if you need to write your own function?

A common method of getting around for loops is the **apply** family of functions. These take a data structure and a function, and applies a function over all the elements in the object.

```{r}
fruit <- c("apple", "orange", "pear", "banana")

# make function that takes in only one element
make.plural <- function(x){
   plural <- paste(x, 's', sep = '') # sep is for collapse, so collpase ''
   return(plural)
}

make.plural('apple')

```

-   `apply()` : loop over the margins (1 = row, 2 = column) of an array
-   `lapply()` : loop over a list then returns a list
-   `sapply()` : loop over a list then returns a named vector
-   `tapply()`: loop over subsets of a vector
-   `mapply()`: multivariate version of `lapply()`. Use this if you have a function that takes in 2 or more arguments.

```{r}
# apply that function to every element
lapply(fruit, make.plural) # returns a list
sapply(fruit, make.plural) # returns a named vector

library(purrr) # load package
map(fruit, make.plural) # type consistent
```

```{r}
# Why sapply is bad 

sapply(1:100, paste) # return character 

sapply(integer(), paste) # return list!

library(purrr)
map(1:100, paste) # return list

map(integer(), paste) # return list
```
