---
title: "How to Write Better Code"
author: "Jae Yeon Kim"
institute: "UC Berkeley"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
name: toc

# Objectives 

- "Your taste develops faster than your skills." - From Jenny Bryan's talk on [Code Smells and Feels](https://github.com/jennybc/code-smells-and-feels/blob/master/2018-07_user-brisbane-bryan.pdf)

- This section is for helping you to develop your taste so that you could avoid smelly code. We're doing this in four steps.

  - Step1: How to write readable code 
  
  - Step2: How to write reusable code 
  
  - Step3: How to create a reproducible question 
  
  - Step4: How to test your code systematically

---

# The sign of bad code (smelly code)

1. Unreadable (you need to ask the author)

2. Not modifiable (if you try to fix a bug or improve a feature, it will take forever.)

>> "Code smell" is an evocative term for that vague feeling of unease we get when reading certain bits of code. It's not necessarily wrong, but neither is it obviously correct. We may be reluctant to work on such code, because past experience suggests it's going to be fiddly and bug-prone. In contrast, there's another type of code that just feels good to read and work on. What's the difference? If we can be more precise about code smells and feels, we can be intentional about writing code that is easier and more pleasant to work on. I've been fortunate to spend the last couple years embedded in a group of developers working on the tidyverse and r-lib packages. Based on this experience, I'll talk about specific code smells and deodorizing strategies for R. - Jenny Bryan

---

# Writing readable code 

> Every major open-source project has its own style guide: a set of conventions (sometimes arbitrary) about how to write code for that project. It is much easier to understand a large codebase when all the code in it is in a consistent style. - Google Style Guides

R also has its recommended style. For instance, see the following examples. 

- [Google's R style guide](https://google.github.io/styleguide/Rguide.xml)

- [R code style guide](http://r-pkgs.had.co.nz/r.html) by Hadley Wickham 

- [The tidyverse style guide](http://style.tidyverse.org/) by Hadley Wickham

---

# Naming matters 

When naming files, remember the following three rules:

- Machine readable (avoid spaces, punctuation, periods, and any other special characters except _ and -)

- Human readable (should be meaningful. No text1, image1, etc.,)

- Ordering (e.g., 01, 02, 03, … )

```{r eval = FALSE}
# Good
fit_models.R

# Bad
fit models.R
```

---

# Naming matters 

When naming objects:

  - Don’t use special characters.

  - Don’t capitalize.

```{r eval = FALSE}

# Good 
day_one
    
# Bad 
DayOne

```

---

# Naming matters 

When naming functions:
  - Don’t use special characters.
  - Don’t capitalize.
  - Use verbs instead of nouns. (Functions do something!)

```{r eval = FALSE}
# Good 
run_rdd # regression discontinuity design

# Bad 
rdd
```

---

# Spacing
 

```{r eval = FALSE}

# Good 

x[, 1] 

mean(x, na.rm = TRUE) 

# Bad

x[,1]

mean (x, na.rm = TRUE)

```

---

# Identing 

Indent at least 4 spaces. 

```{r eval = FALSE}
# Good
if (y < 0) {
  message("y is negative")
}

# Bad
if (y < 0) {
message("Y is negative")}

# Good
do_something_very_complicated(
  something = "that",
  requires = many,
  arguments = "some of which may be long"
)

# Bad
do_something_very_complicated("that", requires = many, arguments =
                              "some of which may be long"
                              )
```

---

# Comments 

- Use comments to explain your decisions.
- But, show your code; Do not try to explain your code by comments.
- Also, try to comment out rather than delete the code that you experiment with.

```{r eval = FALSE}
# Average sleep hours of Jae
jae %>%
  # By week
  group_by(week) %>%
  # Mean sleep hours 
  summarise(week_sleep = mean(sleep, na.rm = TRUE))
```

---

# Pipes 

Pipes (%) = Chaining commands 

```{r eval = FALSE}
# Good
iris %>%
  group_by(Species) %>%
  summarize_if(is.numeric, mean) %>%
  ungroup() %>%
  gather(measure, value, -Species) %>%
  arrange(value)

# Bad
iris %>% group_by(Species) %>% summarize_all(mean) %>%
ungroup %>% gather(measure, value, -Species) %>%
arrange(value)
```

---

# Writing reusable code 

Stop (excessive) copying and pasting. 

>> Copy-and-paste programming, sometimes referred to as just pasting, is the production of highly repetitive computer programming code, as produced by copy and paste operations. It is primarily a pejorative term; those who use the term are often implying a lack of programming competence. It may also be the result of technology limitations (e.g., an insufficiently expressive development environment) as subroutines or libraries would normally be used instead. However, there are occasions when copy-and-paste programming is considered acceptable or necessary, such as for boilerplate, loop unrolling (when not supported automatically by the compiler), or certain programming idioms, and it is supported by some source code editors in the form of snippets. - Wikipedia

It’s okay for pasting for the first attempt to solve a problem. But if you copy and paste three times (a.k.a. Rule of Three in programming), something’s wrong. You’re working too hard. You need to be lazy!

---

# When you need refactoring 

* Let’s imagine `df` is a survey dataset.

* `a, b, c, d` = Survey questions

* `-99`: non-responses

* Your goal: replace `-99` with `NA`

# Data

```{r}
set.seed(1234) # for reproducibility 

df <- data.frame("a" = sample(c(-99, 1:3), size = 5 , replace = TRUE),
             "b" = sample(c(-99, 1:3), size = 5 , replace = TRUE),
             "c" = sample(c(-99, 1:3), size = 5 , replace = TRUE),
             "d" = sample(c(-99, 1:3), size = 5 , replace = TRUE))
```

---

# Copy and paste solution

```{r}
df$a[df$a == -99] <- NA
df$b[df$b == -99] <- NA
df$c[df$c == -99] <- NA
df$d[df$d == -99] <- NA

df
```

---

# Using a function

```{r}
# Create a custom function
fix_missing <- function(x) { # INPUT
  x[x == -99] <- NA # COMPUTATION
  x # OUTPUT 
}

# Apply the function to each column (vector)
# This iterated part can and should be automated.
df$a <- fix_missing(df$a)
df$b <- fix_missing(df$b)
df$c <- fix_missing(df$c)
df$d <- fix_missing(df$d)
```

---

# Automation

```{r}
if (!require(purrr)) install.packages("purrr")

df <- purrr::map_df(df, fix_missing) # What is this magic? We will unpack the blackbox (`map_df()`) later.

df
```

---

# Takeaways 

* Your code becomes more reusable, when it’s easier to change, debug, and scale up. Don’t repeat yourself and embrace the power of lazy programming.

>> The tools and processes that come out of this endeavor fired by laziness will speed up the production. - Philipp Lenssen

* Only when your code becomes reusable, you would become efficient in your data work. Otherwise, you need to start from scratch or copy and paste, when you work on a new project.

>> The key idea in reuse is that parts of a computer program written at one time can be or should be used in the construction of other programs written at a later time. - Wikipedia

---

# Asking a reproducible question 

* A minimal reproducible example (MRE)

1. A minimal dataset
2. The minimal burnable code (i.e., the code that can reproduce the error)
3. The necessary information on package, R version, system (use `sessionInfo()`)
4. A seed for reproducibility (`set.seed()`), if you used a random process.

---

# Asking a reproducible question

Use `reprex` package to create 

```{r}
if (!require(reprex)) install.packages("reprex")
```

Copy the following code and type `reprex()` in the console.

```{r eval = FALSE}
gpa <- c(3, 4, 4, 3)
mean(gpa)
```

---

# Testing systematically

Strongly recommend switching from adhoc testing to formal automated testing (i.e., unit testing)

> Whenever you are tempted to type something into a print statement or a debugger expression, write it as a test instead. — Martin Fowler

```{r}

if (!require(testthat)) install.packages("testthat")

pacman::p_load(testthat)

context("Variable check")

test_that("Check whether instructor variable is setup correctly", {
  
  instructors <- c("Jae", "Nick")

  expect_equal(class(instructors), "character")

}
)

```

---

# Another example 

Inspired by an example in Hadley Wickham's R Journal paper (2011): https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf

```{r}

context("Model check")

test_that("Check whether the model is lm", {
  
  model <- lm(mpg ~ wt, data = mtcars)
  
  # Passes
  expect_that(model, is_a("lm"))

  # Fails
  expect_that(model, is_a("glm"))

})
```

---

# Run tests 

```{r eval = FALSE}

test_file(file.choose()) # file 

test_dir() # directory

auto_test() # the test code tested when you save file 

```



